Notes on Perturbational Quantities

### 2008-07-28 docs/pert_notes/paper.tex
new doc sorting out how rho, rho', (rho X), X, (rho h), (rho h)', h, h', T, and etarho relate to each other depending on predict_rho and enthalpy_pred_type

\def\half  {\frac{1}{2}}
\def\dt    {\Delta t}
\def\kth   {k_{\rm th}}

\def\edge  {\rm EDGE}
\def\mac   {\rm MAC}
\def\trans {\rm TRANS}

\def\eb    {{\bf e}}
\def\fb    {{\bf f}}
\def\ib    {{\bf i}}
\def\Ub    {{\bf U}}

\def\Ubt   {\widetilde{\bf U}}
\def\ut    {\tilde{u}}
\def\vt    {\tilde{v}}
\def\wt    {\tilde{w}}

\title{Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}}

@@

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.  However, this term is present in the cell-update.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?  Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
\item NOTE: The {\tt probin} parameter ``use\_new\_godunov'' enables all four of these features.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
\item {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$ in the computation of $\Ub^{\mac}$, so therefore the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.  Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
\item {\tt MAESTRO} does not corner couple for 3D problems.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

The evolution equation for the perturbational velocity is:
\begin{equation}
\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Perturbational Velocity Equation}
\end{equation}
We are going to use a 1D Taylor series extrapolation in space and time to compute time-centered normal components of $\Ubt^{\trans}$ at face-centers.  By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.  We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.
\subsection{2D Cartesian Case}
For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (x,r)$.\\

For the x-faces, the 1D extrapolation of $u$ is:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
u_{R,\ib-\half\eb_x}^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
\end{eqnarray}
We pick the final edge states using a Riemann solver:
\begin{equation}
u^{\trans}_{\ib-\half\eb_x} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
u_{R,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} < 0, \\
\end{cases}
\end{equation}
For the r-faces, we choose include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.  The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.  The term proportional to $\wt$ is an additional term.  For the r-faces, the 1D extrapolation of $\wt$ is:
\begin{eqnarray}
\wt_{L,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
\wt_{R,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
\end{eqnarray}
The total radial velocity is defined as:
\begin{equation}
w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.\label{Total Radial Velocity}
\end{equation}
The spatial derivatives of $w_0$ are given by the two-point centered difference:
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.\label{Radial Derivative of w_0}
\end{equation}
We pick the final edge states using:
\begin{equation}
\wt^{\trans}_{\ib-\half\eb_r} =
\begin{cases}
0, & \left(w_L^{\trans} \le 0 ~~ {\rm AND} ~~ w_R^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
\wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
\wt_{R,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
\end{cases}
\end{equation}
where $w_L^{\trans} = \wt_{L,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_{R,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
\subsection{3D Cartesian Case}
For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (x,y,r)$.\\

The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.  The procedure is analogous to computing $u^{\trans}$.  For y-faces, the 1D extrapolation of $v$ is:
\begin{eqnarray}
v_{L,\ib-\half\eb_y}^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
v_{R,\ib-\half\eb_y}^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
\end{eqnarray}
\begin{equation}
v^{\trans}_{\ib-\half\eb_y} =
\begin{cases}
0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~~ {\rm AND} ~~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
v_{R,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} < 0. \\
\end{cases}
\end{equation}
\subsection{3D Spherical Case}
For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (x,y,z)$.\\

We first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.  This is a spherical generalization of equation (\ref{Total Radial Velocity}).  We also map $\partial w_0/\partial r$ onto a Cartesian array.  This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).  Then we use a 1D extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).  For example, for x-faces:
\begin{eqnarray}
\ut_L^{\trans} &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
\ut_R^{\trans} &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
\end{eqnarray}
Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.  To compute $\vt^{\trans}$ (or $\wt^{\trans}$), simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).  Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
\cleardoublepage
\section{Computing $\Ubt^{\mac,*}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\

The force terms are:
\begin{equation}
\fb =
\left(\begin{array}{c}
f_u \\
f_w \\
\end{array}\right)
= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
\end{equation}
\subsection{2D Cartesian Case}
We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
\end{eqnarray}
Then we upwind based on $w^{\trans}$:
\begin{equation}
u_{\ib-\half\eb_r}^{1D} =
\begin{cases}
\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
\end{cases}
\end{equation}
Then, we do a full-dimensional extrapolation of $u$ to x-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac,*} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
u_{R,\ib-\half\eb_x}^{\mac,*} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n \nonumber \\
&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
\end{eqnarray}
Finally, we solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac,*} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
\end{cases}
\end{equation}

To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.  We compute a time-centered 1D extrapolation of $\wt$ to x-faces:
\begin{eqnarray}
\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\trans}$:
\begin{equation}
\wt_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
Then, we do a full-dimensional extrapolation of $\wt$ to r-faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
\begin{eqnarray}
\wt_{L,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}) + \frac{\dt}{2}f_{w,\ib-\eb_r},\label{wt_L Extrapolation} \\
\wt_{R,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n \nonumber \\
&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}) + \frac{\dt}{2}f_{w,\ib}.\label{wt_R Extrapolation}
\end{eqnarray}
Next, we solve a Riemann problem:
\begin{equation}
\wt_{\ib-\half\eb_r}^{\mac,*} =
\begin{cases}
0, & \left(w_L^{\mac,*} \le 0 ~~ {\rm AND} ~~ w_R^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 
\end{cases}
\end{equation}
We then project $\Ubt^{\mac,*}$ to obtain $\Ubt^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
%\subsection{3D Cartesian Case}
%\subsection{3D Spherical Case}
\cleardoublepage
\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.  The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.  We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:
\begin{itemize}
\item For {\tt enthalpy\_pred\_type} = 1, we predict $(\rho h)'$ to edges.
\item For {\tt enthalpy\_pred\_type} = 2, we predict $h$ to edges.
\item For {\tt enthalpy\_pred\_type} = 3 and 4, we predict $T$ to edges.
\end{itemize}
The force terms are:
\begin{eqnarray}
f_{\rho'} &=& -\rho'\nabla\cdot\Ub^{\mac} - \nabla\cdot(\rho_0\Ubt^{\mac}), \\
f_{X_k} &=& 0, \\
f_{(\rho h)'} &=& (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} -(\rho h)'\nabla\cdot\Ub^{\mac} - \nabla\cdot\left[(\rho h)_0\Ubt^{\mac}\right] + \nabla\cdot\kth\nabla T \\
f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
\end{eqnarray}
If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.  Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
\subsection{2D Cartesian Case}
We begin by computing a time-centered 1D extrapolation of $s$ to all faces.  The extrapolation of $s$ to x-faces is:
\begin{eqnarray}
s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \\
s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\mac}$:
\begin{equation}
s_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
The extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
\begin{eqnarray}
s_{L,\ib-\half\eb_r}^{1D} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n, \\
s_{R,\ib-\half\eb_r}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n.
\end{eqnarray}
Then we upwind based on $w^{\mac}$:
\begin{equation}
s_{\ib-\half\eb_r}^{1D} =
\begin{cases}
\half\left(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
u_{R,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} < 0. \\
\end{cases}
\end{equation}
Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of $s$ to x-faces is:
\begin{eqnarray}
s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
\end{eqnarray}
Then, we upwind based on $u^{\mac}$.
\begin{equation}
s_{\ib-\half\eb_x}^{\edge} =
\begin{cases}
\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
\end{cases}
\end{equation}
The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
\begin{eqnarray}
s_{L,\ib-\half\eb_r}^{\edge} = s_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n \nonumber \\
&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_r},  \\
s_{R,\ib-\half\eb_r}^{\edge} = s_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n \nonumber \\
&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}. 
\end{eqnarray}
Then, we upwind based on $w^{\mac}$.
\begin{equation}
s_{\ib-\half\eb_r}^{\edge} =
\begin{cases}
\half\left(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
s_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
\end{cases}
\end{equation}
%\subsection{3D Cartesian Case}
%\subsection{3D Spherical Case}
\cleardoublepage
\section{Computing $\Ubt^{\edge}$ in {\tt MAESTRO}}
In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
\subsection{2D Cartesian Case}
We begin by computing a time-centered 1D extrapolation of $u$ to all faces.  The extrapolation of $u$ to x-faces is:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\trans}$:
\begin{equation}
u_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
u_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
The extrapolation of $u$ to r-faces is:
\begin{eqnarray}
u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
\end{eqnarray}
Then we upwind based on $w^{\trans}$:
\begin{equation}
u_{\ib-\half\eb_r}^{1D} =
\begin{cases}
\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
\end{cases}
\end{equation}
We repeat this procedure to compute a time-centered 1D extrapolation of $\wt$ at each face.  The extrapolation of $\wt$ to x-faces is:
\begin{eqnarray}
\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\trans}$:
\begin{equation}
\wt_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
The extrapolation of $\wt$ to r-faces is:
\begin{eqnarray}
\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}. \label{1D wt_R extrap}
\end{eqnarray}
Then we upwind based on $w^{\trans}$:
\begin{equation}
\wt_{\ib-\half\eb_r}^{1D} =

>>>>

\title{Notes on Perturbational Quantities}

@@
